The value of intermediation

### 2023-04-05 src/paper/theory.tex
Switch to one line per sentence

A common feature of the aforementioned papers is that they involve models with a finite number of agents. However, as is often the case in economics, infinite-player approximations might turn out to be more tractable while retaining the main conclusions of the finite models from which they arise. Indeed, in the case of cooperative games with non-atomic players, both the axiomatic approach to defining a value on infinite games and the asymptotic approach to calculate the limit of its finite counterparts lead to the same result \footnote{On the appropriate subspace of non-atomic games.} \parencite{aumann2015values}, which is furthermore analytically simple and readily interpretable. An economic application of this theory can be found in \textcite{billera1978internal}.

Beyond games with no atoms, economists are often interested in situations characterized by a relatively small number of large players and numerous individually rather insignificant ones. The continuous analogue of them are called oceanic games, in which the set of players consists of a finite number of atoms and an ocean made up of a continuum of fringe players. While the axiomatic characterization of the value on such games has not been particularly fruitful,\footnote{\textcite{hart1973values} shows that the usual axioms do not characterize a unique value in general.} \textcite{fogelman1980asymptotic} demonstrates that an asymptotic approach works well on a large subset of them. More specifically, the limit of the players' values converge to relatively simple and easy-to-interpret expressions as the number of small players goes towards infinity.

@@

Consider a cooperative game with two types of players: a major player $P$ and $n$ smaller players. I.e., $N = \{P, A_1, \dots, A_n\}$. Assume that (1) no coalition of players can achieve a positive value without the participation of $P$ and (2) players $A_i$ are identical. Let $n_T(S)$ denote the number of players of type $T \in \{P, A\}$ in coalition $S$. Then, the value that any coalition $S \subset N$ can achieve is the following:
\begin{align*}
    v(S) = \begin{cases}
        0                              & \text{if } P \notin S \\
        f\left(\frac{n_A(S)}{n}\right) & \text{otherwise}.
    \end{cases}
\end{align*}
One can think of this setting as a (one-sided) platform and a number of potential entrants, if the entrants can only reach their prospective consumers through the platform.\footnote{While these assumptions might seem somewhat restrictive, they are relatively simple to relax.} First, let us characterize the conditions under which this game is superadditive. These features provide some theoretical support for using Shapley-values as bargaining outcomes, as papers creating a link between non-cooperative bargaining and cooperative values often rely on such assumptions.\footnote{For example, the results in \textcite{gul1989bargaining} rely on superadditivity, while monotonicity suffices for \textcite[]{hart1996bargaining}.}

@@

In the following, let us assume that the assumptions required for monotonicity are always satisfied. In addition to the previous advantages, it also ensures that the games are of bounded variation, and thus the main theorem in \textcite{fogelman1980asymptotic} applies them, as well. Namely, the asymptotic value coincides with the axiomatic one using the uniform distribution, providing multiple ways of characterizing them. In practice, it means that there is no need to explicitly calculate the limits of finite-player cooperative games, as there is a more direct option for characterizing the value of the oceanic game. Nevertheless, in many of the following theorems, the asymptotic limit is preferred, as I believe that makes the proofs more instructive.
%\textcolor{red}{Note: I realize that this paragraph is somewhat vague and needs explanation. However, for the current set of results, I do not rely on said theorem. If I end up using it later, I will expand on this part.}

@@

\begin{corollary}
    \label{cor:fringe_value}
    The aggregated Shapley-value of the fringe is
    \begin{align*}
        \varphi_A^\infty = f(1) - \int_0^1 f(t) \dt.
    \end{align*}
    Furthermore, if $f$ is differentiable on $[0, 1]$, Then the value of the fringe can also be expressed as
    \begin{align*}
        \varphi_A^\infty = \int_0^1 t f'(t) \dt.
    \end{align*}
\end{corollary}

@@

While the values of oceanic games (i.e. the limits of the Shapley values of their finite approximations in finite games) are well known in the relevant literature, there are no results about the limits of the weighted (Shapley) values. While providing very general results in the form of \textcite[]{fogelman1980asymptotic} is outside of the scope of the paper\footnote{They show that the (within some constraints) any discretization of an oceanic game converges to the same value.} the results in this section still provide some new insights into weighted values of oceanic games. In particular, I show that the weighted value of a game with one large and many small players converges to a rather simple and intuitive expression, if the size of the small players go to zero at the same rate.

Imagine a general non-cooperative game $(N, v)$ with $N = \{1, \dots, n\}$. Assume that, in addition to the value function $v : 2^n \to \mathbb{R}$, the game is also endowed with a weight system $\lambda \in \mathbb{R}^n$. These weights can be thought of as measures of innate bargaining power \parencite{shapley1953additive}, not reflected in the value function. Further support for this interpretation can be found in \textcite{hart1996bargaining}, demonstrating that in a certain alternating offer bargaining game, weights are related to the probability of each player making the offer.\footnote{For example, only one player having a positive value corresponds to them making take it or leave it offers.}

@@

Finally, the following corollaries support the interpretation of weights as bargaining power, as a higher weight corresponds to a higher Shapley-value in this game.\footnote{This would not necessarily have to be the case, as demonstrated by \textcite{owen1968communications}.} Furthermore, the limit of the weighted value of $P$, as their weight goes to zero (infinity) is precisely the payoffs $P$ would achieve if $P$ were making (facing) take-it-or-leave-it offers.

@@

\section{Application}

@@

While the previous procedure might look somewhat peculiar at first glance, its elements would not be out of place in more traditional (non-cooperative) models, either. Principally, the two main ideas it relies are the following. First, bargaining outcomes should not depend on which player actually obtains the money, all that matters is the total that a given set of players can obtain.\footnote{For example, imagine the following two situations with two players. In the first one, the first player obtains a playoff of two dollars if they both decide to cooperate and nothing otherwise. In the second scenario, both players receive one dollar in the case of cooperation. If they can transfer monay between each other, bargaining outcomes should be the same in both scenarios.} Second, players during the bargaining foresee the (unique) subgame perfect equilibrium in the following subgame, and negotiate based on them.

@@

There is a variety of ways through which this paper could be expanded or built upon. On the theoretical side, some of the stricter assumptions could be relaxed, such as fringe agents achieving a zero payoff without the platform. Besides, the convergence result for the weighted values could be strengthened if one could manage to show that any sequence along ever more refined partitions of the atomic players converges to the same limit \parencite[à la][]{fogelman1980asymptotic}. The possibilities for practical applications are no less interesting, either. As a first step, the majorly simplified example application could be expanded upon, so that it is a better description of certain real-life paltform markets. Finally, the cooperative approach to profit distribution could become an alternative for the more usual, but in some sense more extreme assumptions about the bargaining process widely used in industrial organization models.

% Next steps:
% \begin{itemize}
%     \item Show how the equilibrium number of entrants depends on $N_P$
%     \item Show how consumer welfare (which is proportional to $\log(N_f + N_P)$) depends on $N_P$
%     \item Try to characterize the profit-maximizing $N_P$ and how it depends on the parameters
% \end{itemize}

>>>>

A common feature of the aforementioned papers is that they involve models with a finite number of agents.
However, as is often the case in economics, infinite-player approximations might turn out to be more tractable while retaining the main conclusions of the finite models from which they arise.
Indeed, in the case of cooperative games with non-atomic players, both the axiomatic approach to defining a value on infinite games and the asymptotic approach to calculate the limit of its finite counterparts lead to the same result \footnote{
    On the appropriate subspace of non-atomic games.
} \parencite{aumann2015values}, which is furthermore analytically simple and readily interpretable. An economic application of this theory can be found in \textcite{billera1978internal}.

Beyond games with no atoms, economists are often interested in situations characterized by a relatively small number of large players and numerous individually rather insignificant ones.
The continuous analogue of them are called oceanic games, in which the set of players consists of a finite number of atoms and an ocean made up of a continuum of fringe players.
While the axiomatic characterization of the value on such games has not been particularly fruitful,\footnote{
    \textcite{hart1973values} shows that the usual axioms do not characterize a unique value in general.
} \textcite{fogelman1980asymptotic} demonstrates that an asymptotic approach works well on a large subset of them.
More specifically, the limit of the players' values converge to relatively simple and easy-to-interpret expressions as the number of small players goes towards infinity.

@@

Consider a cooperative game with two types of players: a major player $P$ and $n$ smaller players.
I.e., $N = \{P, A_1, \dots, A_n\}$.
Assume that (1) no coalition of players can achieve a positive value without the participation of $P$ and (2) players $A_i$ are identical.
Let $n_T(S)$ denote the number of players of type $T \in \{P, A\}$ in coalition $S$.
Then, the value that any coalition $S \subset N$ can achieve is the following:
\begin{align*}
    v(S) = \begin{cases}
        0                              & \text{if } P \notin S \\
        f\left(\frac{n_A(S)}{n}\right) & \text{otherwise}.
    \end{cases}
\end{align*}
One can think of this setting as a (one-sided) platform and a number of potential entrants, if the entrants can only reach their prospective consumers through the platform.\footnote{
    While these assumptions might seem somewhat restrictive, they are relatively simple to relax.
}
First, let us characterize the conditions under which this game is superadditive.
These features provide some theoretical support for using Shapley-values as bargaining outcomes, as papers creating a link between non-cooperative bargaining and cooperative values often rely on such assumptions.\footnote{
    For example, the results in \textcite{gul1989bargaining} rely on superadditivity, while monotonicity suffices for \textcite[]{hart1996bargaining}.
}

@@

In the following, let us assume that the assumptions required for monotonicity are always satisfied.
In addition to the previous advantages, it also ensures that the games are of bounded variation, and thus the main theorem in \textcite{fogelman1980asymptotic} applies them, as well.
Namely, the asymptotic value coincides with the axiomatic one using the uniform distribution, providing multiple ways of characterizing them.
In practice, it means that there is no need to explicitly calculate the limits of finite-player cooperative games, as there is a more direct option for characterizing the value of the oceanic game.
Nevertheless, in many of the following theorems, the asymptotic limit is preferred, as I believe that makes the proofs more instructive.

@@

\begin{corollary}
    \label{cor:fringe_value}
    The aggregated Shapley-value of the fringe is
    \begin{align*}
        \varphi_A^\infty = f(1) - \int_0^1 f(t) \dt.
    \end{align*}
    Furthermore, if $f$ is differentiable on $[0, 1]$, then the value of the fringe can also be expressed as
    \begin{align*}
        \varphi_A^\infty = \int_0^1 t f'(t) \dt.
    \end{align*}
\end{corollary}

@@

While the values of oceanic games (i.e. the limits of the Shapley values of their finite approximations in finite games) are well known in the relevant literature, there are no results about the limits of the weighted (Shapley) values.
While providing very general results in the form of \textcite[]{fogelman1980asymptotic} is outside the scope of the paper\footnote{
    They show that the (within some constraints) any discretization of an oceanic game converges to the same value.
} the results in this section still provide some new insights into weighted values of oceanic games.
In particular, I show that the weighted value of a game with one large and many small players converges to a rather simple and intuitive expression, if the size of the small players go to zero at the same rate.

Imagine a general non-cooperative game $(N, v)$ with $N = \{1, \dots, n\}$.
Assume that, in addition to the value function $v : 2^n \to \mathbb{R}$, the game is also endowed with a weight system $\lambda \in \mathbb{R}^n$.
These weights can be thought of as measures of innate bargaining power \parencite{shapley1953additive}, not reflected in the value function.
Further support for this interpretation can be found in \textcite{hart1996bargaining}, demonstrating that in a certain alternating offer bargaining game, weights are related to the probability of each player making the offer.\footnote{
    For example, only one player having a positive value corresponds to them making take it or leave it offers.
}

@@

Finally, the following corollaries support the interpretation of weights as bargaining power, as a higher weight corresponds to a higher Shapley-value in this game.\footnote{
    This would not necessarily have to be the case, as demonstrated by \textcite{owen1968communications}.
}
Furthermore, the limit of the weighted value of $P$, as their weight goes to zero (infinity) is precisely the payoffs $P$ would achieve if $P$ were making (facing) take-it-or-leave-it offers.

@@

\section{Example Application}

\section{Old example application}

@@

While the previous procedure might look somewhat peculiar at first glance, its elements would not be out of place in more traditional (non-cooperative) models, either.
Principally, the two main ideas it relies are the following.
First, bargaining outcomes should not depend on which player actually obtains the money, all that matters is the total that a given set of players can obtain.\footnote{
    For example, imagine the following two situations with two players.
    In the first one, the first player obtains a playoff of two dollars if they both decide to cooperate and nothing otherwise.
    In the second scenario, both players receive one dollar in the case of cooperation.
    If they can transfer monay between each other, bargaining outcomes should be the same in both scenarios.
} Second, players during the bargaining foresee the (unique) subgame perfect equilibrium in the following subgame, and negotiate based on them.

@@

There is a variety of ways through which this paper could be expanded or built upon.
On the theoretical side, some of the stricter assumptions could be relaxed, such as fringe agents achieving a zero payoff without the platform.
Besides, the convergence result for the weighted values could be strengthened if one could manage to show that any sequence along ever more refined partitions of the atomic players converges to the same limit \parencite[à la][]{fogelman1980asymptotic}.
The possibilities for practical applications are no less interesting, either.
As a first step, the majorly simplified example application could be expanded upon, so that it is a better description of certain real-life pLAtform markets.
Finally, the cooperative approach to profit distribution could become an alternative for the more usual, but in some sense more extreme assumptions about the bargaining process widely used in industrial organization models.
